Seniority and Hierarchy Configuration Interaction for Radicals and \\ Excited States

### 2023-10-25 Manuscript/manuscript.tex
Delta + update table

\fk{
We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene,
which connects the two equivalent rectangular $D_{2h}$ equilibrium geometries through the square $D_{4h}$ transition state geometry. \cite{Bally_1980,Tallarico_1996}
This is a well-known and challenging problem, requiring high-level calculations to achieve quantitative values for the height of the barrier (see Ref.~\onlinecite{Monino_2022} and references therein).
Here we employ the geometries presented in Ref.~\onlinecite{Monino_2022}, the 6-31+G(d) basis set, and the frozen-core approximation.
We optimize the orbitals for two closed-shell determinants in the $D_{2h}$ geometry, and for two spin-flipped open-shell determinants in the $D_{4h}$ geometry,
from which the excited determinants were generated in the CI calculations.
For this basis set, the reference value of \SI{7.51}{kcal/mol} was obtained at CC with single, double, triple, and quadruple excitations. \cite{Monino_2022}
With respect to this reference value, the errors for the automerization barrier are
\SI{+3.55}{kcal/mol} with our two-determinant calculations, \SI{-2.28}{kcal/mol} with hCI2, \SI{-1.07}{kcal/mol} with hCI2+EN2, \SI{+0.61}{kcal/mol} with CISD, and \SI{-0.49}{kcal/mol} with CISD+EN2.
These values are comparable to those obtained with other methods, extensively discussed in Ref.~\onlinecite{Monino_2022}.
Here we just mention the errors of CC with singles and doubles, \SI{+0.80}{kcal/mol}, and of spin-flip equation-of-motion CC (EOM-CC) with singles and doubles, \SI{-1.65}{kcal/mol},
which share the same $\order*{N^6}$ computational scaling as CISD and hCI2.
This specific example serves to illustrate that the accuracy of hCI models can be similar to that of other methods.
As such, hCI can be taken into consideration in future systematic studies that tackle challenging chemical problems where strong correlation comes into play.
}

>>>>

\fk{
We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene,
which connects the two equivalent rectangular $D_{2h}$ equilibrium geometries through the square $D_{4h}$ transition state geometry. \cite{Bally_1980,Tallarico_1996}
This is a well-known and challenging problem, requiring high-level calculations to achieve quantitative values for the height of the barrier (see Ref.~\onlinecite{Monino_2022} and references therein).
Here we employ the geometries presented in Ref.~\onlinecite{Monino_2022}, the 6-31+G(d) basis set, and the frozen-core approximation.
We optimize the orbitals for two closed-shell determinants in the $D_{2h}$ geometry, and for two spin-flipped open-shell determinants in the $D_{4h}$ geometry,
from which the excited determinants were generated in the subsequent CI calculations.
Even though the calculations target the ground state, they are labelled $\Delta$CI because different references are employed for each geometry.
For this basis set, the reference value of \SI{7.51}{kcal/mol} was obtained at CC with single, double, triple, and quadruple excitations. \cite{Monino_2022}
With respect to this reference value, the errors for the automerization barrier are
\SI{+3.55}{kcal/mol} with our two-determinant calculations, \SI{-2.28}{kcal/mol} with $\Delta$hCI2, \SI{-1.07}{kcal/mol} with $\Delta$hCI2+EN2, 
\SI{+0.61}{kcal/mol} with $\Delta$CISD, and \SI{-0.49}{kcal/mol} with $\Delta$CISD+EN2.
These values are comparable to those obtained with other methods, extensively discussed in Ref.~\onlinecite{Monino_2022}.
Here we just mention the errors of CC with singles and doubles, \SI{+0.80}{kcal/mol}, and of spin-flip equation-of-motion CC (EOM-CC) with singles and doubles, \SI{-1.65}{kcal/mol},
which share the same $\order*{N^6}$ computational scaling as $\Delta$CISD and $\Delta$hCI2.
This specific example serves to illustrate that the accuracy of hCI models can be similar to that of other methods.
As such, hCI can be taken into consideration in future systematic studies that tackle challenging chemical problems where strong correlation comes into play.
}
